### 2019-05-28 Literature_Review/pages/review_content.tex
Changes; - Added a few more references - Adjusted structure of lit' review

\section{Possible research questions}

Overall the research will aim to investigate the strength of pEp's Trustword fingerprint mappings, and the ease in which partial collisions can be obtained for keys and how this will ultimately affect the end user(s).

Key areas that will be looked into:
\begin{itemize}
    \item Is the recommended minimum number of Trustwords enough to provide a basic level of security.
    \item What attributes make a strong general wordlist for fingerprint mapping and does the Trustword implementation exhibit these features.
    \item How easy is it to generate similar keys that attack a targeted key pair?
    \item How can similarity be quantified in terms of words? Does this include pronunciation or visual aspects?
    \item As usability is the main justification for the use of Trustwords are there alternatives that provide the same usability but with more security?
\end{itemize}

\section{Review}

\subsection{Fingerprint representation and comparison}

@@

The first work in this area was performed by Hsu-Chun Hsiao, \textit{et. al.} in their paper: \textit{A Study of User-Friendly Hash Comparison Schemes}\cite{hsiao2009study} in around 2009. \\
The aim of the study was to provide the first insight into the best encoding scheme used to represent a hash fingerprint. The schemes used were Base32, English words, Random Art\footnote{A scheme proposed by A. Perrig \textit{et al.}\cite{perrig1999hash}}, Flag\footnote{Encoding scheme proposed by C Ellison \textit{et al.}\cite{ellison2003public}}, T-Flag\footnote{Yue-Hsun Lin \textit{et al.}\cite{lin2010spate} improvement on Flag}, Flag Ext\footnote{The author's own improvment on Flag} and finally Chinese, Korean and Japanese symbol encoding. 436 participants were assessed in the study.

Alongside the main aim of the paper, the authors wanted to provide empirical links between participant attributes such as age or gender to performance with the encoding scheme. This is something rarely seen in the related research. Hsu-Chun Hsiao \textit{et al.} also provided two categories of similarities; ``hard'' and ``easy''. These were designed to be the worst and best case respectively. As a final aim and the reason for the inclusion of Chinese, Korean and Japanese characters was to research if being a native speaker assists the user in distinguishing subtle differences between encodings in the respective language. The entropy of encodings ranged from 22-bits to 28-bits. This is extremely low in comparison to similar research. However, this can be attributed to the age of the paper. \\
The authors' decision to exclude hexadecimal and numerical encodings (some of the most widespread encoding schemes) was due to their similarity to Base32 and their well-known deficiencies. This is not consistent with available research. For example, in \textit{``Empirical Study of Textual Key-Fingerprint Representations''}\cite{dechand2016empirical} it was shown numerical representations performed significantly better than that of Base32.

Findings from the paper showed that age and gender do not affect the accuracy of the scheme, however, younger participants were significantly faster. \\
The paper recommends Base32, Random Art, T-Flag or their own improvement; Flag Ext. These encoding recommendations were also supplemented with a review of the requirements required by these schemes. The recommendation of graphical encoding schemes is inconsistent with alternative literature, where most other papers have found graphical encodings easy to use but insecure.
In terms of language comprehension; being able to speak the language assists in the ability to discern the differences between hard pairs. Subsequently, knowledge of the language did not help with easy pairs.

@@

Further research bu Kainda \textit{et al.}\cite{kainda2009usability} in 2009 investigated fingerprint representations and comparison methods in the context of using humans as the Out-Of-Band (OOB) channel for secure device pairing. This study was comprehensive and assessed a wide variety of concepts. This is one of the few papers in the literature that assessed the way to compare encodings alongside the representations themselves. For comparison methods the paper looked into ``Compare-and-Confirm'', ``Compare-and-Select'', ``Compare-and-Enter'' and ``Barcode'' scanning. This set includes the most common ways of comparison according to available research. Alongside this, the paper reviews Numerical, Alphanumerical, Words, Sentences, Images, Melodies and Sound (Numerical/Alphanumerical). Again, this in comparison to similar literature is highly comprehensive.

In terms of empirical data gathered by the paper, they reviewed the time completed to asses the representation and the number of security and non-security related errors. Furthermore, the authors also collected user opinion on ease-of-use, satisfaction and confidence of the reviewed schemes. The overall schemes were then quantified using a single metric known as the Single Usability Metric (SUM)\cite{sauro2005method}. This provides a quick way to rank and compare the schemes, thus improving the usability of the results. 
However, the one apparent downside to the experimental side of the study is the 40 participants enrolled. This is one of the smallest sizes of enrolment in the literature and thus, puts a possible limit on the accuracy and applicability of the obtained results.

Overall, if usability is the only consideration the paper ranked the comparison methods: Compare-and-Confirm, Compare-and-Enter, Compare-and-Select and Barcode. If, however, the consideration of security alongside usability is required the best is: Compare-and-Enter, Barcode, Compare-and-Confirm and Compare-and-Select. Research has been limited in this area of modes of comparison, however, the results from this paper are consistent with the alternative literature.
\\
In terms of encoding schemes, the paper ranked Numerical, Alphanumerical and Words as the top three encoding schemes (These were all with Compare-and-Confirm). This took into consideration the usability and security of the scheme. This is relatively consistent with other literature where numerical and written encoding schemes seem to be the most effective overall.

@@

A paper by Ersin Uzun \textit{et al.}\cite{uzun2007usability} exclusively reviewed comparison methods. This is performed in the context of ``Secure device pairing'' and thus some comparison schemes are tested that are not relevant in a fingerprint comparison context, for example  ``Choose-and-Enter'' (A participant chooses a passphrase from an available list and enters it into the other device). Thus, only relevant comparison scheme and their results have been extracted and compared.

The relevant comparison methods tested were, Compare-and-Confirm, 
Compare-and-Select and Compare-and-Enter. These are almost exactly the same encoding schemes assessed in the relevant research literature. The paper also had two rounds of experiments including 40 participants each where changes were made to UX and format between these studies. 
Due to the ``Secure device pairing'' context, consideration needs to be made into the actual strings being encoded. Due to secure device pairing commonly utilising Short Authentication Strings (SAS) the
paper only reviewed encodings of 4 digit long strings. This, therefore,limits the applicability of the results in comparison to other fingerprint encoding literature.

Between the two rounds, the authors dropped ``Confirm-and-Enter'' due to its low usability and over similarity to a similar scheme called ``Copy'' (used for passphrase verification). The low rating for Confirm-and-Enter is in contrast with research performed by Kainda \textit{et al.}\cite{kainda2009usability} They ranked Confirm-and-Enter as the 2nd most usable system.

With changes to the UX of Confirm-and-Confirm backed up by external research \cite{palmer1990attentional}\cite{hammer2009category} the second round of research was performed with a new set of participants. The alterations took Compare-and-Confirm from 20\% to 0\% fatal error rate (FER). The same trend for Confirm-and-Select with it reducing its FER from 12.5\% to 5\%. The authors, therefore, recommended ``Compare-and-Confirm'' as the overall best choice for comparison methods. Again, comparing this word to the most similar work done by Kainda \textit{et al.}\cite{kainda2009usability} this is consistent with their findings where they ranked ``Compare-and-Confirm'' as the most secure comparison method.

@@

Work by Dechand \textit{et al.}\cite{dechand2016empirical} empirically investigated the usability of 4 distinct textual representations evaluated with an experiment involving a total of 1047 participants. The textual representations were: Alphanumeric, Numeric, Words and Sentences. They assessed the number of attacks missed with each scheme alongside results from a questionnaire on the participants preferred scheme and perceived usability.

The paper touches upon issues with decentralised methods of identification such a PGP's Web of Trust and the problems these solutions have with user adoption. These points are made in an attempt to validate the requirement for manual comparison of key based fingerprints. This is a common theme that appears in the majority of the reviewed papers. Other references made within the paper touch upon the vulnerabilities and usability issues present with the way humans interact with the security systems. Examples of these are studies showing humans find it difficult to comprehend long and ``meaningless" strings and the lack of actual comparison performed by users in live scenarios. These are, however, acknowledged as limitations in the later stages of the paper.

The paper has defined the upper and lower bound costs of the attacker's resources and strength as \$610K to \$16B, with an ability to control 80-bits of the fingerprint. This in comparison to other papers is high and is almost encroaching into the realm of a highly sophisticated attacker. Therefore, the lack of consideration for the lower-end of the attack resource spectrum can be considered another limitation.

@@

J Tan \textit{et al.}\cite{tan2017can} work is very similar to that of the research already discussed. Eight distinct fingerprint representations were tested with over 661 participants. Each representation was tested using ``Compare-and-Confirm'' and ``Compare-and-Select''. The small difference with this paper is the inclusion of two novel graphical encodings Vash\footnote{https://github.com/thevash/vash} and Unicorns\footnote{https://unicornify.pictures/}. The compared schemes were: Hexadecimal, Alternating vowel/consonants, Words, Numbers, Sentences, OpenSSH visual host key, Vash and Unicorns.

@@

Further work in the context of secure messaging pairing was produced by M. Shirvanian \textit{et al.}\cite{shirvanian2017pitfalls} in 2017. This paper was unique for a number of reasons. First was to consideration for ``remote-vs-proximity'' pairing where, this is the first consideration of this aspect found in the literature. Second, was their studies aim into investigating fingerprint security and usability in a end-to-end encryption context. The study compared Numerical, Visual and Auditory representations. This is a standard setup for similar studies in the literature. QR code verification was also added to the proximity tests. They split the variance of attack severity into three levels; one character change, one block change and changes to the entire fingerprint.

@@

In conclusion for this section, there has been extensive work in the investigation of the usability of textual and visual encoding schemes alongside methods of comparison. There are mixed opinions about visual representations, however, the majority defines them as easy to use but severely lacking in security. Research on textual representations, however, is in agreement; natural language based encodings such as words or sentences provide some of the highest usability and security over the baseline performance of alphanumerical encodings. The literature also concurs on the use of ``Compare-and-Confirm'' as the best method of comparing fingerprints overall, with Compare-and-Select being highlighted for its poor security and usability.\\
The starkest limitations of the literature are the range of participant and encoding entropy, where at its worst only 22-bits of entropy were tested. Another limitation is the scant amount of research solely investigating modes of comparison, where the literature only contains two papers with this aim. Even these limitations, as a whole the literature on this topic has provided a relatively comprehensive and coherent view on the usability of textual and visual hash encoding schemes.

@@

Some of the oldest preliminary work into visual encoding schemes was performed by Adrian Perrig \textit{et al}\cite{perrig1999hash}. in the creation of their scheme ``Random Art" in 1999. The motivation for creating such a scheme was the perceived flaws in the ways humans verify and compare written information. As mentioned in previous sections visual encoding schemes have been shown to have mixed success, with low security being one of their most alarming flaws. This research laid the foundation for further work in analysing the security of visual encoding schemes (as discussed in the previous section).

Further research into the creation of unique visual hash schemes have been performed by C Ellison \textit{et al.} \cite{ellison2003public} (Flag), Yue-Hsun Lin \textit{et al.}\cite{lin2010spate} (T-Flag) and work by MM Olembo \textit{et al.}\cite{olembo2013developing}. Each publication has provided a new way to visually represent a key fingerprint. Alongside the academic literature, there are more informally presented methods of visual fingerprints such as Unicorns\footnote{https://unicornify.pictures/} and Robots\footnote{https://github.com/e1ven/Robohash}. This list is by no means exhaustive but is used to depict the amount of research and work invested into graphical hash representations.

One paper of note is the preliminary work performed by D Loss \textit{et al.}\cite{loss2009drunken} in their \textit{``An analysis of the OpenSSH fingerprint visualization algorithm''} where their aim was to spur on further research with their initial findings into the security of the OpenSSH scheme. The authors claim that the use of the algorithm in OpenSSH is only heuristically defined and there is a need for a formal proof of its security. \\
The paper proposed a number of ways to generate similar fingerprints. The methods proposed were: Naive brute force, Graph Theory, and brute force of a full visual set. They were only able to produce only very basic results and have proposed a large amount of potential further work. Since the paper's publication in 2009, there seems to have been no research building on the work of the authors. This highlights the current gap in the available literature.

Minimal research has also focused on basic textual fingerprint representations and their respective security. Work by A. Karole and N. Saxena\cite{karole2009improving} looked into ways to improve the security of a textual representation. This research aim is to improve the secure device pairing process of comparing two numerical values. The devices used (Nokia 6030b; Mid-range devices at the time of publication) and short SAS compared results in findings that are not directly applicable in a fingerprint comparison context.

A more specific subsection of textual fingerprints is the use of words and sentences to encode hash digests. Some of the first work in this area was produced by Patrick Juola and Philip Zimmermann 
\cite{juola1996whole} and their work in generating a word list where phonetic distinctiveness was prioritised. Each word is mapped to one byte where the unique aspect of the word list is the separation of ``even'' and ``odd'' words where ``even'' byte positions are sample from the even-list and ``odd'' from the odd-list. This effectively creates two sub-word lists. The pairing of these word lists were determined through the use of a Genetic algorithm with phonetic distance as the metric. The paper also includes a study on effective measures of ``linguistic distances'' of words and provided an in-depth discussion into these areas.

@@

Other research of note is work by M. Goodrich \textit{et al.}\cite{goodrich2006loud} called \textit{Loud and Clear: Human-Verifiable Authentication Based on Audio}. As the name suggests the authors were researching ways to improve current methods of secure device pairing. The unique aspect of this work is the use of a Text-to-Speech system reading out syntactically correct English sentences. The sentences are based on a MadLibs\footnote{https://en.wikipedia.org/wiki/Mad\_Libs} where static placeholders were replaced with potential words.\\
The work into a potential wordlist can be seen as an extension to the work performed by Juola and Zimmermann\cite{juola1996whole} as they aimed to emulate the techniques used in PGPfone. The paper's finding are limited by the lack of empirical data backing up claims made by the author as the systems performance and security are only theoretically assessed.

Aside from this research, there have been further informal implementations of fingerprint encodings. The first being by Michael Rogers\footnote{https://github.com/akwizgran/basic-english}. Rogers' implementation is a program designed to map fingerprints to pseudo-random poems. This implementation was again, empirically evaluated by Dechand \textit{et al.}\cite{dechand2016empirical}. Older work by N. Haller with the S/KEY\cite{haller1995s} shows the implementation of a system designed to represent a hash as a series of six short words. However, this system is designed for a one-time-password purpose and only provides word mappings for basic human usability of the password and not within a fingerprint verification context. Therefore, the wordlist has not been designed with pronounceability in mind.
\\
A very recent implementation of a word list can be found in Pretty Easy Privacy (pEp) implementation of TrustWords\footnote{https://tools.ietf.org/html/draft-birk-pep-trustwords-03}. The unique aspect of TrustWords is its mapping of a single word to 16-bits. In comparison to other literature, this is the highest number of bits-per-word seen. Full mappings (no duplication of words) would, therefore, require $2^{16}$ words in the dictionary and arguably is higher than most users vocabulary. This deviation from the norm has not been currently backed up by research. Therefore, highlighting on a gap in the current literature.

@@

\section{Attacks on fingerprint verification}
This area of research studies ways to physically execute attacks on fingerprint encoding schemes. This differs from previously examined work due to papers discussing the performance and fallibility of encoding schemes simulated the attack without consideration for how the attack would be performed.

@@

Research in 2002 by Konrad Rieck\cite{rieck2002fuzzy} is the first formalisation of attacks on fingerprint representations. The paper titled \textit{``Fuzzy Fingerprints Attacking Vulnerabilities in the Human Brain''}
aimed to look into ways users check hexadecimal encoded OpenSSH fingerprint representations. The author created an elegant way to `weight' more important chunks of the digest. The bytes furthest to the right and left of the digests provided the highest weight. The weight was the smallest in the centre of the digest. This provides a way to score digests and determine the best partial collisions found. For example with the target fingerprint: \verb|9F23| a partial match \verb|9313| is given a score of 45\% even though only two characters were matching. This is due to the weightings.

@@

The only other relevant research on this topic is the work by M 
Shirvanian \textit{et al.}\cite{shirvanian2014wiretapping} 
and their paper \textit{``Wiretapping via Mimicry: Short 
Voice Imitation Man-in-the-Middle Attacks on Crypto 
Phones''}. In this paper, the authors develop a way to 
impersonate users when authenticating 
Short-authentication-Strings (SAS) in pairing of 
Crypto-phones. To achieve this impersonating they propose 
two methods: ``Short voice reordering attack'' where an 
arbitrary SAS string is recreated by re-ordering snipped 
obtained from eavesdropping a previous connection
and ``Short voice morphing attacks'' whereby the use of 
previously eavesdropped audio snippets the attacker can
morph their own voice to match that of the victim. With 
these methods, they aimed to attack encodings of Numbers, 
PGP word list (previously discussed work by Juola and 
Zimmermann \cite{juola1996whole}) and MadLib (M. Goodrich 
\textit{et al.}\cite{goodrich2006loud} work also 
previously discussed). The effectiveness of these attacks 
were evaluated with a study involving 30 participants.

>>>>

% \section{Possible research questions}

% Overall the research will aim to investigate the strength of pEp's Trustword fingerprint mappings, and the ease in which partial collisions can be obtained for keys and how this will ultimately affect the end user(s).

% Key areas that will be looked into:
% \begin{itemize}
%     \item Is the recommended minimum number of Trustwords enough to provide a basic level of security.
%     \item What attributes make a strong general wordlist for fingerprint mapping and does the Trustword implementation exhibit these features.
%     \item How easy is it to generate similar keys that attack a targeted key pair?
%     \item How can similarity be quantified in terms of words? Does this include pronunciation or visual aspects?
%     \item As usability is the main justification for the use of Trustwords are there alternatives that provide the same usability but with more security?
% \end{itemize}

\section{Review - Fingerprint representations}

\subsection{Fingerprint representation comparison}

@@

\textbf{Hsu-Chun Hsiao, \textit{et. al.}} produced the first work in this area in their paper: \textit{A Study of User-Friendly Hash Comparison Schemes}\cite{hsiao2009study} in around 2009. The aim of the study was to provide the first insight into the best encoding scheme used to represent a hash fingerprint. The schemes used were Base32, English words, Random Art\cite{perrig1999hash}, Flag\cite{ellison2003public}, T-Flag\cite{lin2010spate}, Flag Ext\footnote{The author's own improvement on Flag} and finally Chinese, Korean and Japanese symbol encoding. A total of 436 participants were assessed in the study.

Alongside the main aim of the paper, the authors wanted to provide empirical links between participant attributes (such as age or gender) and performance with the encoding scheme. This is something rarely seen in the related research. Hsu-Chun Hsiao \textit{et al.} also provided two categories of similarities; ``hard'' and ``easy''. These were designed to be the worst and best case respectively. As a final aim and the reason for the inclusion of Chinese, Korean and Japanese characters was to research if being a native speaker assists the user in distinguishing subtle differences between encodings in the respective language. The entropy of encodings ranged from 22-bits to 28-bits. This is extremely low in comparison to similar research. However, this can be attributed to the age of the paper. \\
The authors' decision to exclude hexadecimal and numerical encodings (some of the most widespread encoding schemes) was due to their similarity to Base32 and their well-known deficiencies. This is not consistent with available research. For example, in \textit{``Empirical Study of Textual Key-Fingerprint Representations''}\cite{dechand2016empirical} it was shown numerical representations performed significantly better than that of Base32.

Findings from the paper showed that age and gender do not affect the accuracy of the scheme, however, younger participants were significantly faster. \\
The paper recommends Base32, Random Art, T-Flag or their own improvement; Flag Ext. These encoding recommendations were also supplemented with a review of the requirements required by these schemes. The recommendation of graphical encoding schemes is inconsistent with alternative literature, where most other papers have found graphical encodings easy to use but insecure.
In terms of language comprehension; being able to speak the language assists in the ability to discern the differences between hard pairs. Subsequently, knowledge of the language did not assist in differentiating easy pairs as these had high accuracy regardless of the spoken language.

@@

\textbf{Kainda \textit{et al.}}\cite{kainda2009usability} in 2009 investigated fingerprint representations and comparison methods in the context of using humans as the Out-Of-Band (OOB) channel for secure device pairing. This study was comprehensive and assessed a wide variety of concepts. This is one of the few papers in the literature that assessed the way to compare encodings alongside the representations themselves. For comparison methods the paper looked into ``Compare-and-Confirm'', ``Compare-and-Select'', ``Compare-and-Enter'' and ``Barcode'' scanning. This set includes the most common ways of comparison according to available research. Alongside this, the paper reviews Numerical, Alphanumerical, Words, Sentences, Images, Melodies and Sound (Numerical/Alphanumerical). Again, this in comparison to similar literature is highly comprehensive.

In terms of empirical data gathered by the paper, they reviewed the time completed to asses the representation and the number of security and non-security related errors. Furthermore, the authors also collected user opinion on ease-of-use, satisfaction and confidence of the reviewed schemes. The overall schemes were then quantified using a metric known as the Single Usability Metric (SUM)\cite{sauro2005method}. This provides an elementary way to rank and compare the schemes, thus improving the usability of the results. 
However, the one apparent downside to the experimental side of the study is the 40 participants enrolled. This is one of the smallest sizes of enrolment in the literature and thus, puts a possible limit on the accuracy and applicability of the obtained results.

Overall, if usability is the only consideration the paper ranked the comparison methods: Compare-and-Confirm, Compare-and-Enter, Compare-and-Select and Barcode. If, however, the consideration of security alongside usability is required the best is: Compare-and-Enter, Barcode, Compare-and-Confirm and Compare-and-Select. Research has been limited in this area of modes of comparison, however, the results from this paper are consistent with the limited alternative literature. In terms of encoding schemes, the paper ranked Numerical, Alphanumerical and Words as the top three encoding schemes (These were all with Compare-and-Confirm). This took into consideration the usability and security of the scheme. This is relatively consistent with other literature where numerical and written encoding schemes seem to be the most effective overall.

@@

\textbf{Ersin Uzun \textit{et al.}}\cite{uzun2007usability} exclusively reviewed comparison methods. This is performed in the context of ``Secure device pairing'' and thus some comparison schemes are tested that are not relevant in a fingerprint comparison context, for example  ``Choose-and-Enter'' (A participant chooses a passphrase from an available list and enters it into the other device). Thus, only relevant comparison scheme and their results have been extracted and compared.

The relevant comparison methods tested were,Compare-and-Confirm, Compare-and-Select and Compare-and-Enter. These are almost exactly the same encoding schemes assessed in the relevant research literature. The paper also had two rounds of experiments including 40 participants each where changes were made to UX and format between these studies. 
Due to the ``Secure device pairing'' context, consideration needs to be made into the actual strings being encoded. Due to secure device pairing commonly utilising Short Authentication Strings (SAS) the
paper only reviewed encodings of 4 digit long strings. This, therefore,limits the applicability of the results in comparison to other fingerprint encoding literature. Between the two rounds, the authors dropped ``Confirm-and-Enter'' due to its low usability and over similarity to a similar scheme called ``Copy'' (used for passphrase verification). The low rating for Confirm-and-Enter is in contrast with research performed by Kainda \textit{et al.}\cite{kainda2009usability} They ranked Confirm-and-Enter as the 2nd most usable system. With changes to the UX of Confirm-and-Confirm backed up by external research \cite{palmer1990attentional}\cite{hammer2009category} the second round of research was performed with a new set of participants. The alterations took Compare-and-Confirm from 20\% to 0\% fatal error rate (FER). The same trend for Confirm-and-Select with it reducing its FER from 12.5\% to 5\%. The authors, therefore, recommended ``Compare-and-Confirm'' as the overall best choice for comparison methods. Again, comparing this word to the most similar work done by Kainda \textit{et al.}\cite{kainda2009usability} this is consistent with their findings where they ranked ``Compare-and-Confirm'' as the most secure comparison method.

@@

\textbf{Dechand \textit{et al.}}\cite{dechand2016empirical} empirically investigated the usability of 4 distinct textual representations evaluated with an experiment involving a total of 1047 participants. The textual representations were: Alphanumeric, Numeric, Words and Sentences. They assessed the number of attacks missed with each scheme alongside results from a questionnaire on the participants preferred scheme and perceived usability.

The paper touches upon issues with decentralised methods of identification such a PGP's Web of Trust and the problems these solutions have with user adoption. These points are made in an attempt to validate the requirement for manual comparison of key based fingerprints. This is a common theme that appears in the majority of the reviewed papers. Other references made within the paper touch upon the vulnerabilities and usability issues present with the way humans interact with the security systems. Examples of these are studies showing humans find it difficult to comprehend long and ``meaningless" strings and the lack of actual comparison performed by users in live scenarios. These are, however, acknowledged as limitations in the later stages of the paper. The paper has defined the upper and lower bound costs of the attacker's resources and strength as \$610K to \$16B, with an ability to control 80-bits of the fingerprint. This in comparison to other papers is high and is almost encroaching into the realm of a highly sophisticated attacker. Therefore, the lack of consideration for the lower-end of the attack resource spectrum can be considered another limitation.

@@

\textbf{J Tan \textit{et al.}}\cite{tan2017can} work is very similar to that of the research already discussed. Eight distinct fingerprint representations were tested with over 661 participants. Each representation was tested using ``Compare-and-Confirm'' and ``Compare-and-Select''. The small difference with this paper is the inclusion of two novel graphical encodings Vash\footnote{https://github.com/thevash/vash} and Unicorns\footnote{https://unicornify.pictures/}. The compared schemes were: Hexadecimal, Alternating vowel/consonants, Words, Numbers, Sentences, OpenSSH visual host key, Vash and Unicorns.

@@

\textbf{M. Shirvanian \textit{et al.}}\cite{shirvanian2017pitfalls} produced further work in the context of secure messaging pairing. This paper was unique for a number of reasons. First was to consideration for ``remote-vs-proximity'' pairing where, this is the first consideration of this aspect found in the literature. Second, was their studies aim into investigating fingerprint security and usability in a end-to-end encryption context. The study compared Numerical, Visual and Auditory representations. This is a standard setup for similar studies in the literature. QR code verification was also added to the proximity tests. They split the variance of attack severity into three levels; one character change, one block change and changes to the entire fingerprint.

@@

In conclusion for this section, there has been extensive work in the investigation of the usability of textual and visual encoding schemes alongside methods of comparison. There are mixed opinions about visual representations, however, the majority defines them as easy to use but severely lacking in security. Research on textual representations, however, is in agreement; natural language based encodings such as words or sentences provide some of the highest usability and security over the baseline performance of alphanumerical encodings. The literature also concurs on the use of Compare-and-Confirm as the best method of comparing fingerprints overall, with Compare-and-Select being highlighted for its poor security and usability.\\
The starkest limitations of the literature are the range of participant and encoding entropy, where at its worst only 22-bits of entropy were tested. Another limitation is the scant amount of research solely investigating modes of comparison, where the literature only contains two papers with this aim. Even these limitations, as a whole the literature on this topic has provided a relatively comprehensive and coherent view on the usability of textual and visual hash encoding schemes.

@@

Some of the oldest preliminary work into visual encoding schemes was performed by \textbf{Adrian Perrig \textit{et al}}\cite{perrig1999hash}. in the creation of their scheme ``Random Art" in 1999. The motivation for creating such a scheme was the perceived flaws in the ways humans verify and compare written information. As mentioned in previous sections visual encoding schemes have been shown to have mixed success, with low security being one of their most alarming flaws. This research laid the foundation for further work in analysing the security of visual encoding schemes (as discussed in the previous section).

Further research into the creation of unique visual hash schemes have been performed by \textbf{C Ellison \textit{et al.}} \cite{ellison2003public} (Flag), \textbf{Yue-Hsun Lin \textit{et al.}}\cite{lin2010spate} (T-Flag) and work by \textbf{M.  Olembo \textit{et al.}}\cite{olembo2013developing}. Each publication has provided a new way to visually represent a key fingerprint. Alongside the academic literature, there are more informally presented methods of visual fingerprints such as Unicorns\footnote{https://unicornify.pictures/} and Robots\footnote{https://github.com/e1ven/Robohash}. This list is by no means exhaustive but is used to depict the amount of research and work invested into graphical hash representations.

One paper of note is the preliminary work performed by \textbf{D Loss \textit{et al.}}\cite{loss2009drunken} in their \textit{``An analysis of the OpenSSH fingerprint visualization algorithm''} where their aim was to spur on further research with their initial findings into the security of the OpenSSH scheme. The authors claim that the use of the algorithm in OpenSSH is only heuristically defined and there is a need for a formal proof of its security. \\
The paper proposed a number of ways to generate similar fingerprints. The methods proposed were: Naive brute force, Graph Theory, and brute force of a full visual set. They were only able to produce only very basic results and have proposed a large amount of potential further work. Since the paper's publication in 2009, there seems to have been no research building on the work of the authors. This highlights the current gap in the available literature.

Minimal research has also focused on basic textual fingerprint representations and their respective security. Work by \textbf{A. Karole} and \textbf{N. Saxena}\cite{karole2009improving} looked into ways to improve the security of a textual representation. This research aim is to improve the secure device pairing process of comparing two numerical values. The devices used (Nokia 6030b; Mid-range devices at the time of publication) and short SAS compared results in findings that are not directly applicable in a fingerprint comparison context.

A more specific subsection of textual fingerprints is the use of words and sentences to encode hash digests. Some of the first work in this area was produced by \textbf{Juola} and \textbf{Zimmermann}
\cite{juola1996whole} and their work in generating a word list where phonetic distinctiveness was prioritised. Each word is mapped to one byte where the unique aspect of the word list is the separation of ``even'' and ``odd'' words where ``even'' byte positions are sample from the even-list and ``odd'' from the odd-list. This effectively creates two sub-word lists. The pairing of these word lists were determined through the use of a Genetic algorithm with phonetic distance as the metric. The paper also includes a study on effective measures of ``linguistic distances'' of words and provided an in-depth discussion into these areas.

@@

Other research of note is work by \textbf{M. Goodrich \textit{et al.}}\cite{goodrich2006loud} called \textit{Loud and Clear: Human-Verifiable Authentication Based on Audio}. As the name suggests the authors were researching ways to improve current methods of secure device pairing. The unique aspect of this work is the use of a Text-to-Speech system reading out syntactically correct English sentences. The sentences are based on a MadLibs\footnote{https://en.wikipedia.org/wiki/Mad\_Libs} where static placeholders were replaced with potential words.\\
The work into a potential wordlist can be seen as an extension to the work performed by Juola and Zimmermann\cite{juola1996whole} as they aimed to emulate the techniques used in PGPfone. The paper's finding are limited by the lack of empirical data backing up claims made by the author as the systems performance and security are only theoretically assessed.

Aside from this research, there have been further informal implementations of fingerprint encodings. The first being by \textbf{Michael Rogers}\footnote{https://github.com/akwizgran/basic-english}. Rogers' implementation is a program designed to map fingerprints to pseudo-random poems. This implementation was again, empirically evaluated by Dechand \textit{et al.}\cite{dechand2016empirical}. Older work by N. Haller with the S/KEY\cite{haller1995s} shows the implementation of a system designed to represent a hash as a series of six short words. However, this system is designed for a one-time-password purpose and only provides word mappings for basic human usability of the password and not within a fingerprint verification context. Therefore, the wordlist has not been designed with pronounceability in mind.
\\
A very recent implementation of a word list can be found in Pretty Easy Privacy (pEp) implementation of TrustWords\footnote{https://tools.ietf.org/html/draft-birk-pep-trustwords-03}. The unique aspect of TrustWords is its mapping of a single word to 16-bits. In comparison to other literature, this is the highest number of bits-per-word seen. Full mappings (no duplication of words) would, therefore, require $2^{16}$ words in the dictionary and arguably is higher than most users vocabulary. This deviation from the norm has not been currently backed up by research. Therefore, highlighting on a gap in the current literature.

@@

\subsection{Fingerprint representation attacks}
This area of research studies ways to physically execute attacks on fingerprint encoding schemes. This differs from previously examined work due to papers discussing the performance and fallibility of encoding schemes simulated the attack without consideration for how the attack would be performed.

@@

Research in 2002 by \textbf{Konrad Rieck}\cite{rieck2002fuzzy} is the first formalisation of attacks on fingerprint representations. The paper titled \textit{``Fuzzy Fingerprints Attacking Vulnerabilities in the Human Brain''}
aimed to look into ways users check hexadecimal encoded OpenSSH fingerprint representations. The author created an elegant way to `weight' more important chunks of the digest. The bytes furthest to the right and left of the digests provided the highest weight. The weight was the smallest in the centre of the digest. This provides a way to score digests and determine the best partial collisions found. For example with the target fingerprint: \verb|9F23| a partial match \verb|9313| is given a score of 45\% even though only two characters were matching. This is due to the weightings.

@@

The only other relevant research on this topic is the work by \textbf{M Shirvanian \textit{et al.}}\cite{shirvanian2014wiretapping} 
and their paper \textit{``Wiretapping via Mimicry: Short 
Voice Imitation Man-in-the-Middle Attacks on Crypto 
Phones''}. Further research in the area of ``human voice impersonation'' has received lots of attention \cite{mukhopadhyay2015all}\cite{chen2017you}\cite{wu2015spoofing}. This paper was chosen over other alternatives due to is specific use of encoding schemes in its evaluation.

In this paper, the authors develop a way to 
impersonate users when authenticating 
Short-authentication-Strings (SAS) in pairing of 
Crypto-phones. To achieve this impersonating they propose 
two methods: ``Short voice reordering attack'' where an 
arbitrary SAS string is recreated by re-ordering snipped 
obtained from eavesdropping a previous connection
and ``Short voice morphing attacks'' whereby the use of 
previously eavesdropped audio snippets the attacker can
morph their own voice to match that of the victim. With 
these methods, they aimed to attack encodings of Numbers, 
PGP word list (previously discussed work by Juola and 
Zimmermann \cite{juola1996whole}) and MadLib (M. Goodrich 
\textit{et al.}\cite{goodrich2006loud} work also 
previously discussed). The effectiveness of these attacks 
were evaluated with a study involving 30 participants.
